Real Estate

### 2022-11-13 src/real_estate.Rmd
updating scores for updated list of MSAs

MSA_demo_data <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=0")

@@

MSA_median_rent_data <- read_sheet("https://docs.google.com/spreadsheets/d/1oxNxWQRWpANbgb603ShBlw84QkyvsRaH4TzHKKVWhxw/edit#gid=0")
```

```{r}
# select columns in demographic data
MSA_demographics <- MSA_demo_data %>% 
  clean_names() %>% 
  # convert house prices to be in thousands
  mutate(house_price_2021 = house_price_2021 * 1000) %>% 
  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, house_price_2021)
```

```{r}
# aggregate yardi data into MSAs by cbsa
MSA_realEstate_data <- city_realEstate_data %>% 
  clean_names() %>% 
   # calculate change in rent
  mutate(rent_change = (x2022_rent - average_rent_2021)/average_rent_2021 * 100) %>% 
  group_by(cbsa) %>% 
  # calculate average values for MSAs with more than one city
  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
            avg_rent_2022 = mean(x2022_rent, na.rm = TRUE),
            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
            avg_rent_change = mean(rent_change, na.rm = TRUE))

@@

```{r}
# add MSA_median rent data
MSA_realEstate_compiled<- MSA_median_rent_data %>% 
  select(cbsa, rent50_0) %>%
  mutate(median_rent = as.numeric(rent50_0)) %>% 
  left_join(MSA_realEstate_data) %>% 
  left_join(MSA_demographics)
```

```{r}
# combine real estate and demographic data by cbsa
MSA_RE_ratios <- MSA_realEstate_compiled %>% 
      # calculate income to home price
  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
  # calculate rent to income ratio
  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
  mutate(med_rent_to_income = median_rent/median_income_usd_2021)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- MSA_RE_ratios %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>% 
  # normalize employment growth
  mutate(n_employ_change = (employment_yo_y21_22 - min(employment_yo_y21_22)) / (max(employment_yo_y21_22) - min(employment_yo_y21_22))) %>% 
  # normalize rent to income
  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  mutate(n_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

```{r}
# manually check weighted score
n_MSA_realEstate<- n_MSA_realEstate %>% 
  mutate(weighted_pop_change = 0.35 * n_pop_change) %>% 
  mutate(weight_emp = 0.20 * n_employ_change) %>% 
  mutate(weighted_occ = 0.15 * n_avg_occupancy) %>% 
  mutate(weighted_rent_change = 0.10 * n_rent_change) %>% 
  mutate(weighted_med_rent_income = 0.10 * n_med_rent_to_income_inverted) %>% 
  mutate(weight_income_homeprice = 0.10 * n_income_to_homeprice) %>% 
  mutate(total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_rent_change + weighted_med_rent_income + weight_income_homeprice)
```

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, msa, starts_with("n_"), total_score) %>% 
  mutate(real_estate_score = msa) %>% #this is just a place holder
  mutate(RE_score_avg_rent = msa)
```

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice[i]
                  )
  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
}

>>>>

@@

select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")

@@

```{r}
# aggregate yardi data into MSAs by cbsa
MSA_realEstate_data <- city_realEstate_data %>% 
  clean_names() %>% 
  rename(avg_rent2022 = x2022_rent) %>% 
   # calculate change in rent
  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
  group_by(cbsa) %>% 
  # calculate average values for MSAs with more than one city
  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
            avg_rent_change = mean(rent_change, na.rm = TRUE))

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  # normalize rent change
  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

```{r}
# manually check weighted score
n_MSA_realEstate<- n_MSA_realEstate %>% 
  mutate(weighted_pop_change = 0.35 * n_pop_change) %>% 
  mutate(weight_emp = 0.20 * n_employ_change) %>% 
  mutate(weighted_occ = 0.15 * n_avg_occupancy) %>% 
  mutate(weighted_avg_rent_change = 0.10 * n_avg_rent_change) %>% 
  mutate(weighted_med_rent_change = 10* n_med_rent_change) %>% 
  mutate(weighted_med_rent_income = 0.10 * n_med_rent_to_income_inverted) %>% 
  mutate(weight_income_homeprice = 0.10 * n_income_to_homeprice) %>% 
  mutate(med_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_med_rent_change + weighted_med_rent_income + weight_income_homeprice) %>% 
  mutate(avg_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_avg_rent_change + weighted_avg_rent_income + weight_income_homeprice)
```

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_"), total_score) %>% 
  mutate(real_estate_score_med = msa) %>% #this is just a place holder
  mutate(RE_score_avg_rent = msa)
```

@@

for (i in 1:size[1]) {
  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
                  n_MSA_realEstate$n_employ_change[i],
                  n_MSA_realEstate$n_avg_occupancy[i],
                  n_MSA_realEstate$n_rent_change[i],
                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
                  n_MSA_realEstate$n_income_to_homeprice[i]
                  )
  n_MSA_realEstate$real_estate_score_med[i] = weighted.mean(MSA_x_data, real_estate_weights)
}
